Validity of a model for a cantilever two-dimensional beam

### 2024-03-13 article.tex
Draft article template fix

\documentclass[preprint,11pt]{elsarticle}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{draftwatermark}
\usepackage{tabularx} % Required for tables that stretch to text width
\usepackage{booktabs} % For prettier tables
\usepackage{ltablex}
\usepackage{xcolor,cancel,bm,tikz}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{amsmath,amssymb,amsthm,amsfonts}
\usepackage{enumitem}
\usepackage{kantlipsum,graphicx}
\usepackage{multicol,colortbl,multirow}
\usepackage{pgfplots,booktabs,xfrac,floatrow}
\usepackage{color}

\usepackage{setspace}
\setstretch{0.7} % Adjust this value as needed

\setlength{\parindent}{0pt}

\usepackage{geometry}
\geometry{
  left=1in,
  right=1in,
  top=1in,
  bottom=1in,
}

\SetWatermarkText{DRAFT}
\SetWatermarkScale{1}

\journal{Mathematical and Computer Modelling}

@@

\author{Rudi du Toit}
\address{University of Pretoria, Pretoria, South Africa}

\begin{abstract}
Your abstract here.
\end{abstract}

\begin{keyword}
Vibration \sep Cantilever \sep Beam \sep Three-Dimensional \sep Two-Dimensional
\end{keyword}

@@

\section{Introduction}
Your introduction here.

\section{Models}

\subsection{Three-dimensional model}
Consider a vector valued displacement function $u \in \Omega \subset R^3$.

@@

\subsubsection*{Constitutive Equation}
For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$ as follows.\\

@@

If the principal stresses $\sigma_i$ are all non-zero, then Hooke's law can be written in an alternative form.\\

\textbf{Hooke's law in the alternative form}
\begin{eqnarray}
	T = \left( \frac{E}{1+\nu} \right)\mathcal{E} + \frac{\nu E}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE}.
\end{eqnarray}
With $\textrm{Tr}(\mathcal{E}) = \epsilon_{11} + \epsilon_{12} + \epsilon_{13}$ is the trace of the strain operator $\mathcal{E}$.\\

@@

\subsection{Dimensionless form} \label{ssec:3D_Model:DimensionlessForm}
Suppose $\ell$ represents some notable dimension (e.g.the length) of the elastic body 
and $G$ the shear modulus of elasticity.\\

@@

\textbf{Remark} Some of the constants are introduced to simplify the equations and also allow for comparisons to other models such as the Timoshenko beam model.

@@

\subsection{Model problems}\label{ssec:3D_Model:ModelProblems}
Divide the boundary of $\Omega$ into two distinct parts, referred to as $\Sigma$ and $\Gamma$. The following will be considered a model problem for a three-dimensional elastic body. 
The following will be considered a model problem for a three-dimensional elastic body (also for the two-dimensional elastic body).
\begin{enumerate}
	\item[] The equations of motion \eqref{eq:3D_Model:EM-D} and \eqref{eq:3D_Model:divT-D} is satisfied in $\Omega$;
	\item[] Hooke's law \eqref{eq:3D_Model:CE-D} is satisfied in $\Omega$;
	\item[] The displacement $u = u_\Sigma$ is specified on
		$\Sigma$; The traction $Tn = t_\Gamma$ is specified on
		$\Gamma$.
\end{enumerate} \label{sym:n}\label{sym:SigmaGamma}

\subsubsection{Problem 3D}
Find a vector valued function $u$, satisfying equations
\eqref{eq:3D_Model:EM-D} to \eqref{eq:3D_Model:CE-D} and the following boundary
conditions:\\

@@

\subsection{Variational form}\label{ssec:3D_Model:VariationalForm}
Define the bilinear forms
\begin{eqnarray}
	b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dV \label{eq:3D_Model:Bilinear}
\end{eqnarray}\label{sym:b}
and
\begin{eqnarray}
	c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dV \label{eq:3D_Model:Bilinear_c}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1+\nu)(1-2\nu)}$.\\

@@

The variational problem of Problem 3DV is given as:
\subsubsection{Problem 3DV}\label{sssec:3D_Model:Problem3D1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi), \label{eq:3D_Model:Problem3D1VEq}
\end{align}
for all $\phi \in T(\Omega)$\\

@@

\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\sigma_{11} & = & \frac{1}{\gamma(1-\nu^2)} ( \varepsilon_{11} + \nu\varepsilon_{22})\\
		\sigma_{22} & = & \frac{1}{\gamma(1-\nu^2)} (\nu \varepsilon_{11} + \varepsilon_{22}).
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray}
		\sigma_{12} & = & \frac{1}{\gamma(1+\nu)} \varepsilon_{12} \label{stress_comp}
	\end{eqnarray}
\end{minipage}\\

@@

\subsection{Two-dimensional model}
Using \cite{Sadd05} as a guide, a special case of general plane stress is considered.

@@

It follows that $u_3$ is a constant on $\Omega$ and $u$ is a vector valued function in 
$R^2$ with two-dimensional stress and strain. The out of plane conditions for strain 
\eqref{plane_stress_neseccary_condition} falls away and Hooke’s law can be written in a 
two-dimensional form using the stress components (1.2.16) as 
\[ T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{tr}(\mathcal{E})I.\]

\subsection{Equations of motion and constitutive equations}\label{ssec:2D_Model:EquationsOfMotion+ConstitutiveEquations}
 The following equations of motion and constitutive equations follow from subsections \ref{ssec:3D_Model:DimensionlessForm} and 1.3.2.
\subsubsection{Equation of motion}\label{sssec:2D_Model:EquationOfMotion}
\begin{eqnarray}
	\partial_t^2 u & = & \textrm{div}T + Q, \label{eq:2D_Model:EM}
\end{eqnarray}
where
\begin{eqnarray}
	\textrm{div} T & = &
	\begin{bmatrix}
		\partial_1 \sigma_{11} + \partial_2 \sigma_{12} \\
		\partial_1 \sigma_{21} + \partial_2 \sigma_{22}
	\end{bmatrix}. \label{eq:2D_Model:DivT}
\end{eqnarray}
\subsubsection{Constitutive equation}\label{sssec:2D_Model:ConstitutiveEquation}
\begin{eqnarray}
	T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{Tr}(\mathcal{E}) I. \label{eq:2D_Model:CE}
\end{eqnarray}

@@

\subsubsection{Problem 2D}\label{sssec:2D_Model:Problem2D1}
Find a vector valued function $u$, satisfying equations \eqref{eq:2D_Model:EM} to \eqref{eq:2D_Model:CE} and the boundary conditions:\\

Boundary Conditions:\\
\begin{eqnarray*}
	u & = & 0 \quad \textrm{ on } \Sigma,\\
	Tn & = & 0 \quad \textrm{ on } \Gamma.
\end{eqnarray*} with $n$ the outward unit normal. This model problem is a cantilever two-dimensional body.

\subsection{Variational form}\label{ssec:2D_Model:VariationalForm}
Define the bilinear forms
\begin{eqnarray}
	b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA \label{eq:2D_Model:Bilinear}
\end{eqnarray}
and
\begin{eqnarray}
	c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dA \label{eq:2D_Model:Bilinear_c}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.\\
Also define the integral
\begin{eqnarray}
	(f,g) &=& \int_{\Omega} f\cdot g \ dA \label{eq:2D_Model:Bilinear_int}
\end{eqnarray}

@@

\subsubsection{Problem 2DV}\label{sssec:2D_Model:Problem2D1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEq}
\end{align}
for all $\phi \in T(\Omega)$.\\

\section{Existance and uniqueness of solutions}
The models in this article concern the vibration of elastic bodies. These type of models
have a similar variational form as the wave equation and are called second order hyperbolic
type problems. The articles \cite{VV02} and \cite{VS19} provide existence and uniqueness
results for these type of problems. In \cite{VV02}, it is assumed that $b$ is symmetric. 
\cite{VS19} extends the results of \cite{VV02} to the case where $b$ is not symmetric. For
the models in the article, \cite{VV02} is a sufficient reference.\\

@@

Consider a general second order hyperbolic type problem GVar. This general problem is used by the 
authors of \cite{VV02} to prove existence and uniqueness results. The models in this article are 
special cases of this general problem.\\

\subsubsection*{Problem GVar}\label{sssec:existence:ProblemGVar}
Given a function $f:J\rightarrow X$, find a function $u\in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$ and for each $t\in J,\ u(t)\in V,\ u'(t) \in V,\ u''(t)\in W$ and
\begin{eqnarray}
	c(u''(t),v)+a(u'(t),v)+b(u(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVar}
\end{eqnarray}
while $u(0)=u_{0},\ \ \ u'(0)=u_{1}$.\\

@@

\section{Modal analysis and comparison of linear models}
Modal analysis is a important concept in the study of vibration of elastic bodies and also 
useful in the comparison of linear models. The following section is a short discussion of 
the basic concepts of modal analysis and how it enables the comparison of linear models.\\

@@

\subsubsection*{Problem GVar}\label{sssec:existence:ProblemGVar}
Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
\begin{eqnarray}
c(u''(t),v)+b(u(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
\end{eqnarray}
with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.\\

@@

Therefore $\displaystyle -\frac{T''(t)}{T(t)}$ must be constant. Suppose that $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$. The existence of such a $\lambda$ is uncertain at this point. So we consider the following eigenvalue problem.\\

@@

\subsection{Validity of series solution}
Consider the following question: When is the formal series solution valid for the vibration problem with initial values $u(0) = u_0$ and $u'(0) = u_1$? To answer this question, we again refer to the article \cite{CVV18}.\\

@@

Therefore $\mathcal{E}(t) = \mathcal{E}(0)$ for all $t>0$. For the case of weak damping, it can be proved that $\mathcal{E}(t) \leq \mathcal{E}(0)$ for all $t>0$ which is the case of \cite{CVV18}.\\

Denote the partial sum $u^{N}(t) = \sum_{n=1}^{N} T_{n}(t)w_n$. Ideally
\begin{eqnarray*}
	u_0^{N} = \sum_{n=1}^{N} T_n(0) w_n \ \ \textrm{ and } \ \ u_{1}^{N} =\sum_{n=1}^{N} T'_n(0) w_n,
\end{eqnarray*}
but $T_n$ is not uniquely defined by the differential equation.\\

Substitution shows that these partial sums with initial conditions $u^N(0) = u^N_0$ and $(u^N)'(0) = u^N_1$ are solutions for \eqref{eq:existence:ProblemGVarHom} in Problem GVar.\\

The authors then define an error function $u^E_N = u - u^N$. Since both $u$ and $u^N$ satisfies \eqref{eq:existence:ProblemGVarHom}, so does this error function with the following the initial conditions, $u^E_N(0) = u_0 - u^N_0$ and $(u^E_N)'(0) = u_1 - u^N_1$.\\

@@

Then $||u_0 - u^N_0||_V^2 \rightarrow 0$ and $||u_1 - u^N_1||_W^2 \rightarrow 0$ as $N \rightarrow \infty$. Therefore $\mathcal{E}(t) \rightarrow 0$ as $N \rightarrow \infty$ by \eqref{eq:inequality}.\\

@@

\subsection{Comparison of models}
The results on modal analysis and the validity of the series solution enable the comparison of different linear models. The solution of the model problems can be expressed as a series solution of the eigenvalues and eigenvectors of the models. Therefore the models can be compared by comparing the eigenvalues and eigenvectors of the models.\\

@@

\section{Finite element method}

\subsection{Three-dimensional model}
Consider a cantilever three-dimensional elastic body, with a rectangular cross-section.

\subsubsection*{Reference configuration for rectangular cross-section}
Let $\{e_1,e_2,e_3\}$ be a right-handed orthonormal basis for $R^3$. Denote the elastic body as $\Omega \in R^3$ with $(0,0,0)$ the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as
\begin{eqnarray*}
	\Omega = \left\{ x \in R^3 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} , \ -\frac{b}{2} \leq x_3 \leq \frac{b}{2}\right \}
\end{eqnarray*} \label{sym:width}
Let $\partial \Omega$ denote the boundary of the body. Divide $\partial \Omega$ into the six distinct flat surfaces as follows:

Using this notation, the boundary conditions are similar to the two-dimensional model in section \ref{sssec:2D_Model:RefConf}, where the body is clamped rigidly at $\Sigma$, and free-hanging on the other sides denoted by $\Gamma$.

\subsubsection{Cantilever elastic body}
Consider a three-dimensional elastic body with rectangular cross-section, rigidly clamped to a surface at attached at the side $\Sigma$ and free-hanging at all the other sides. This is Problem 3D in section \ref{ssec:3D_Model:ModelProblems}.

In section \ref{ssec:3D_Model:VariationalForm}, the variational problem for the three-dimensional cantilever model is defined by Problem 3DV. This general form can now be rewritten as in a model specific form using the reference configuration. For convenience, some of the results from section \ref{sec:3D_Model} are repeated here.

\subsection{Galerkin approximation}
As mentioned in the previous section, section \ref{2d_FEM_G}, the body $\Omega$ needs to be descritised. For this is three-dimensional body, three-dimensional elements shapes are required. The elements used in this dissertation are rectangular prismatic elements. These elements are also known as `brick-shaped' elements and is described in \cite{Wu06}. This shape of element is a natural choice for a three-dimensional elastic body with a rectangular cross-section.

For the this three-dimensional model, piecewise Hermite tri-cubic basis functions are used. Although this is more complex than using piecwise tri-linear basis functions, as mentioned in section \ref{2d_FEM_G}, the tri-cubic basis functions ensure faster convergence and also the derivatives of the solutions.

Consider a cantilever two-dimensional elastic body, with a rectangular cross-section.

@@

\subsubsection{Cantilever elastic body}
Consider a two-dimensional elastic body clamped rigidly to a surface at $x_1 = 0$. The body is free-hanging on the other boundaries. This is Problem 2D in section \ref{ssec:2D_Model:ModelProblem}.

@@

In section \ref{ssec:2D_Model:VariationalForm} the variational problem for the two dimensional cantilever model is derived and refered to as Problem 2DV. Some of the results from section \ref{sec:2D_Model} are repeated here for convenience. Using the reference configuration, these results are rewritten from a general form to a model specific form.

@@

\section{Results}

\section{Validity of a model for a cantilever two-dimensional beam} \label{sec:validity-of-a-2d-beam}

In section \ref{sec:validity-of-a-cantilever-timoshenko-beam}, the article \cite{LVV09} was discussed. In this article, the authors investigated the validity of a cantilever Timoshenko beam by comparing it to a cantilever two-dimensional beam. In this section, the article is extended and the validity of the cantilever two-dimensional model is investigated.

As mentioned in the introduction of the chapter, a beam is a three-dimensional body and therefore a three-dimensional model is more realistic. However the authors of \cite{LVV09} mention that a direct comparison of the one and three-dimensional models will have complexities and they suggest using the two-dimensional model as an intermediate step.

@@

\subsection{The models}
The two-dimensional model is the same model as used in the previous section, Problem T-2 defined in section \ref{ssec:2D_Model:ModelProblem}. From section \ref{ssec:3D_Model:ModelProblems}, the cantilever three-dimensional model, referred to as Problem 3D.

Figure \ref{fig:2Dv3D} shows the two beams side-by-side.

\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{

>>>>

% \documentclass[11pt]{elsarticle}
\input{preamble_article.tex}
% \usepackage{geometry}
% \geometry{
%   left=1in,
%   right=1in,
%   top=1in,
%   bottom=1in,
% }

@@

\author{Rudi du Toit}
\address{Department of Mathematics, University of Pretoria, Pretoria, South Africa}

\journal{Applied Mathematics and Computation}

\begin{abstract}
Beams are wildly used in engineering. The the real world, beams are complex three-dimensional objects. But often, simplified models are used to study the vibration of beams. These simplified models are less complex and computationally less expensive. But the validity of these simplified models is not always clear. In this article, the validity of a two-dimensional model for a cantilever beam is investigated. The two-dimensional model is compared to a three-dimensional model.
\end{abstract}

\begin{keyword}
	Beam \sep Cantilever \sep Two-Dimensional Model \sep Three-Dimensional Model \sep Vibration Problems
\end{keyword}

@@

\section{Introduction}
In the article \cite{LVV09}, the authors investigate the validity of a Timoshenko beam using a two-dimensional model as a reference. The two-dimensional model is a simplified model of the Timoshenko beam. The authors use the Finite Element Method (FEM) to solve the eigenvalue problem for the two models. The results are then compared to determine the validity of the two-dimensional model.

In this article, \cite{LVV09} is extended and the validity of a two-dimensional model is investigated, using a three-dimensional model as a reference. Similar to \cite{LVV09}, the investigation focusses on beam-type problems.

The models are presented in a dimensionless variational form. It is also explained how the two-dimensional model is derived from the three-dimensional model, as a special case of plane stress. The existence and uniqueness of solutions for the two models are briefly discussed.

The concept of modal analysis is then introduced using a general vibration problem. From modal analysis, the formal series solution for the vibration problem is obtained. This formal series solution is a linear combination of the eigenvectors and eigenvalues of the problem. This enables us to compare the solutions of the two models, by only considering a comparison of the eigenvectors and eigenvalues.

Using the Finite Element Method (FEM), a formula for the models are obtained to numerically solve the eigenvalue problem for the two models. The eigenvalues are matched up by comparing the mode shapes, and the eigenvalues are then compared.

\section{Models}
\subsection{Three-dimensional model}
Consider a vector valued displacement function $u \in \Omega \subset R^3$.

@@

\subsubsection*{Constitutive Equation}
For an isotropic material, Hooke's law can be written in terms of Young's modulus $E$ and Poisson's ratio $\nu$ as follows.

@@

If the principal stresses $\sigma_i$ are all non-zero, then Hooke's law can be written in an alternative form.

\textbf{Hooke's law in the alternative form}
\begin{eqnarray}
	T = \left( \frac{E}{1+\nu} \right)\mathcal{E} + \frac{\nu E}{(1+\nu)(1-2\nu)}\textrm{Tr}(\mathcal{E})I \label{eq:3D_Model:CE}.
\end{eqnarray}
With $\textrm{Tr}(\mathcal{E}) = \epsilon_{11} + \epsilon_{12} + \epsilon_{13}$ is the trace of the strain operator $\mathcal{E}$.

@@

\subsection{Dimensionless form} \label{ssec:3D_Model:DimensionlessForm}
Suppose $\ell$ represents some notable dimension (e.g.the length) of the elastic body 
and $G$ the shear modulus of elasticity.

@@

\textbf{Remark} The constants introduced are the same as those in \cite{LVV09}. They simplify the equations and allow for comparisons to other models such as the Timoshenko beam model, although that is not covered in this article.

@@

\subsection{Model problems}\label{ssec:3D_Model}
Divide the boundary of $\Omega$ into two distinct parts, referred to as $\Sigma$ and $\Gamma$. The following will be considered a model problem for a three-dimensional elastic body. 
The following will be considered a model problem for a three-dimensional elastic body (also for the two-dimensional elastic body).
\begin{enumerate}
	\item[] The equations of motion \eqref{eq:3D_Model:EM-D} and \eqref{eq:3D_Model:divT-D} is satisfied in $\Omega$;
	\item[] Hooke's law \eqref{eq:3D_Model:CE-D} is satisfied in $\Omega$;
	\item[] The displacement $u = u_\Sigma$ is specified on
		$\Sigma$; The traction $Tn = t_\Gamma$ is specified on
		$\Gamma$.
\end{enumerate} \label{sym:n}\label{sym:SigmaGamma}

\subsubsection{Problem 3D}
Find a vector valued function $u$, satisfying equations
\eqref{eq:3D_Model:EM-D} to \eqref{eq:3D_Model:CE-D} and the following boundary
conditions:

@@

\subsection{Variational form}\label{ssec:3D_Model:VariationalForm}
Define the bilinear forms
\begin{eqnarray}
	b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dV \label{eq:3D_Model:Bilinear}
\end{eqnarray}\label{sym:b}
and
\begin{eqnarray}
	c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dV \label{eq:3D_Model:Bilinear_c}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1+\nu)(1-2\nu)}$.

@@

The variational problem of Problem 3DV is given as:
\subsubsection{Problem 3DV}\label{sssec:3D_Model:Problem3D1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi), \label{eq:3D_Model:Problem3D1VEq}
\end{align}
for all $\phi \in T(\Omega)$

@@

\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\sigma_{11} & = & \frac{1}{\gamma(1-\nu^2)} ( \varepsilon_{11} + \nu\varepsilon_{22})\\
		\sigma_{22} & = & \frac{1}{\gamma(1-\nu^2)} (\nu \varepsilon_{11} + \varepsilon_{22}).
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray}
		\sigma_{12} & = & \frac{1}{\gamma(1+\nu)} \varepsilon_{12} \label{stress_comp}
	\end{eqnarray}
\end{minipage}

@@

\subsection{Two-dimensional model} \label{ssec:2D_Model}
For the two-dimensional modek, a special case of general plane stress is considered.

@@

It follows that $u_3$ is a constant on $\Omega$ and $u$ is a vector valued function in 
$R^2$ with two-dimensional stress and strain. The out of plane conditions for strain 
\eqref{plane_stress_neseccary_condition} falls away and Hooke’s law can be written in a 
two-dimensional form using the stress components as 
\[ T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{tr}(\mathcal{E})I.\]

\subsection*{Equations of motion and constitutive equations}\label{ssec:2D_Model:EquationsOfMotion+ConstitutiveEquations}
 The following equations of motion and constitutive equations follow from subsections \ref{ssec:3D_Model:DimensionlessForm} and 1.3.2.
\subsubsection{Equation of motion}\label{sssec:2D_Model:EquationOfMotion}
\begin{eqnarray}
	\partial_t^2 u & = & \textrm{div}T + Q, \label{eq:2D_Model:EM}
\end{eqnarray}
where
\begin{eqnarray}
	\textrm{div} T & = &
	\begin{bmatrix}
		\partial_1 \sigma_{11} + \partial_2 \sigma_{12} \\
		\partial_1 \sigma_{21} + \partial_2 \sigma_{22}
	\end{bmatrix}. \label{eq:2D_Model:DivT}
\end{eqnarray}
\subsubsection*{Constitutive equation}\label{sssec:2D_Model:ConstitutiveEquation}
\begin{eqnarray}
	T = \frac{1}{\gamma(1+\nu)}\mathcal{E} + \frac{\nu}{\gamma(1-\nu^2)}\textrm{Tr}(\mathcal{E}) I. \label{eq:2D_Model:CE}
\end{eqnarray}

@@

\subsubsection*{Problem 2D}\label{sssec:2D_Model:Problem2D1}
Find a vector valued function $u$, satisfying equations \eqref{eq:2D_Model:EM} to \eqref{eq:2D_Model:CE} and the boundary conditions:

Boundary Conditions:
\begin{eqnarray*}
	u & = & 0 \quad \textrm{ on } \Sigma,\\
	Tn & = & 0 \quad \textrm{ on } \Gamma.
\end{eqnarray*} with $n$ the outward unit normal. This model problem is a cantilever two-dimensional body.

\subsection{Variational form}\label{ssec:2D_Model:VariationalForm}
Define the bilinear forms
\begin{eqnarray}
	b(u,\phi) = \int_\Omega~c_1\textrm{Tr}({\cal E}\Phi)+ c_2\textrm{Tr}({\cal E})
	\textrm{Tr}(\Phi) ~dA \label{eq:2D_Model:Bilinear}
\end{eqnarray}
and
\begin{eqnarray}
	c(u,\phi) = \int_\Omega~ (\partial^2_t u) \cdot \phi~dA \label{eq:2D_Model:Bilinear_c}
\end{eqnarray}
with $\displaystyle c_1 = \frac{1}{\gamma(1+\nu)}$ and $\displaystyle c_2 = \frac{\nu}{\gamma(1-\nu^2)}$.
Also define the integral
\begin{eqnarray}
	(f,g) &=& \int_{\Omega} f\cdot g \ dA \label{eq:2D_Model:Bilinear_int}
\end{eqnarray}

@@

\subsubsection{Problem 2DV}\label{sssec:2D_Model:Problem2D1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{eq:2D_Model:Problem2D1VEq}
\end{align}
for all $\phi \in T(\Omega)$.

\section{Existance and uniqueness of solutions}
Our two models, the three-dimensional modela and the two-dimensional model, concern the vibration of elastic bodies. These type of models have a similar variational form as the wave equation and are called second order hyperbolic type problems. The articles \cite{VV02} and \cite{VS18} provide existence and uniqueness results for these type of problems. In \cite{VV02}, it is assumed that $b$ is symmetric. \cite{VS18} extends the results of \cite{VV02} to the case where $b$ is not symmetric. For the models in the article, \cite{VV02} is a sufficient reference.

@@

Consider a general second order hyperbolic type problem GVar. This general problem is used by the 
authors of \cite{VV02} to prove existence and uniqueness results. The models in this article are 
special cases of this general problem.

\subsubsection*{Problem GVar}\label{sssec:existence:ProblemGVar}
Given a function $f:J\rightarrow X$, find a function $u\in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$ and for each $t\in J,\ u(t)\in V,\ u'(t) \in V,\ u''(t)\in W$ and
\begin{eqnarray}
	c(u''(t),v)+a(u'(t),v)+b(u(t),v)= (f(t),v)_{X} \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVar}
\end{eqnarray}
while $u(0)=u_{0},\ \ \ u'(0)=u_{1}$.

@@

\section{Modal analysis and comparison of linear models}
Modal analysis is a important concept in the study of vibration of elastic bodies and also 
useful in the comparison of linear models. The following section is a short discussion of 
the basic concepts of modal analysis and how it enables the comparison of linear models.

@@

\subsubsection*{Problem GVar}\label{sssec:existence:ProblemGVar}
Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
\begin{eqnarray}
c(u''(t),v)+b(u(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
\end{eqnarray}
with initial conditions $u(0) = u_0$ and $u'(0) = u_1$.

@@

Therefore $\displaystyle -\frac{T''(t)}{T(t)}$ must be constant. Suppose that $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$. The existence of such a $\lambda$ is uncertain at this point. So we consider the following eigenvalue problem.

@@

\subsection{Validity of series solution}
Consider the following question: When is the formal series solution valid for the vibration problem with initial values $u(0) = u_0$ and $u'(0) = u_1$? To answer this question, we again refer to the article \cite{CVV18}.

@@

Therefore $\mathcal{E}(t) = \mathcal{E}(0)$ for all $t>0$. For the case of weak damping, it can be proved that $\mathcal{E}(t) \leq \mathcal{E}(0)$ for all $t>0$ which is the case of \cite{CVV18}.

Denote the partial sum $u^{N}(t) = \sum_{n=1}^{N} T_{n}(t)w_n$. Ideally
\begin{eqnarray*}
	u_0^{N} = \sum_{n=1}^{N} T_n(0) w_n \ \ \textrm{ and } \ \ u_{1}^{N} =\sum_{n=1}^{N} T'_n(0) w_n,
\end{eqnarray*}
but $T_n$ is not uniquely defined by the differential equation.

Substitution shows that these partial sums with initial conditions $u^N(0) = u^N_0$ and $(u^N)'(0) = u^N_1$ are solutions for \eqref{eq:existence:ProblemGVarHom} in Problem GVar.

The authors then define an error function $u^E_N = u - u^N$. Since both $u$ and $u^N$ satisfies \eqref{eq:existence:ProblemGVarHom}, so does this error function with the following the initial conditions, $u^E_N(0) = u_0 - u^N_0$ and $(u^E_N)'(0) = u_1 - u^N_1$.

@@

Then $||u_0 - u^N_0||_V^2 \rightarrow 0$ and $||u_1 - u^N_1||_W^2 \rightarrow 0$ as $N \rightarrow \infty$. Therefore $\mathcal{E}(t) \rightarrow 0$ as $N \rightarrow \infty$ by \eqref{eq:inequality}.

@@

\subsection{Comparison of models}
The results on modal analysis and the validity of the series solution enable the comparison of different linear models. The solution of the model problems can be expressed as a series solution of the eigenvalues and eigenvectors of the models. Therefore the models can be compared by comparing the eigenvalues and eigenvectors of the models.

@@

\section{Finite Element Method}
In the section, the Finite Element Method (FEM) is applied to the models to obtain an eigenvalue problem that can be solved numerically.

\subsection{Two-dimensional model}\label{ssec:2D_Model:FEM}
Consider a cantilever two-dimensional elastic body, with a rectangular cross-section.

@@

\subsubsection{Cantilever elastic body}
Consider a two-dimensional elastic body clamped rigidly to a surface at $x_1 = 0$. The body is free-hanging on the other boundaries. This is Problem 2D in section \ref{ssec:2D_Model}.

\FloatBarrier

@@

\FloatBarrier

The variational problem for the two dimensional cantilever model is also derived in \ref{ssec:2D_Model} and refered to as Problem 2DV. Using the reference configuration, the variational problem can be rewritten into a model specific form.

@@

\subsection{Three-dimensional model} \label{ssec:3D_Model:FEM}
Consider a cantilever three-dimensional elastic body, with a rectangular cross-section.

\subsubsection*{Reference configuration for rectangular cross-section}
Let $\{e_1,e_2,e_3\}$ be a right-handed orthonormal basis for $R^3$. Denote the elastic body as $\Omega \in R^3$ with $(0,0,0)$ the point of reference. For a rectangular cross-section, the body $\Omega$ can be described as
\begin{eqnarray*}
	\Omega = \left\{ x \in R^3 \ | \ 0 \leq x_1 \leq 1, \ -\frac{h}{2} \leq x_2 \leq \frac{h}{2} , \ -\frac{b}{2} \leq x_3 \leq \frac{b}{2}\right \}
\end{eqnarray*}
Let $\partial \Omega$ denote the boundary of the body. Divide $\partial \Omega$ into the six distinct flat surfaces as follows:

Using this notation, the boundary conditions are similar to the two-dimensional model, where the body is clamped rigidly at $\Sigma$, and free-hanging on the other sides denoted by $\Gamma$.

\subsubsection{Cantilever elastic body}
Consider a three-dimensional elastic body with rectangular cross-section, rigidly clamped to a surface at attached at the side $\Sigma$ and free-hanging at all the other sides. This is Problem 3D in section \ref{ssec:3D_Model}.

In section \ref{ssec:3D_Model}, the variational problem for the three-dimensional cantilever model is defined by Problem 3DV. This general form can now be rewritten as in a model specific form using the reference configuration.

\subsection{Galerkin approximation}
For this is three-dimensional body, three-dimensional elements shapes are required. The elements used in this dissertation are rectangular prismatic elements. These elements are also known as `brick-shaped' elements and is described in \cite{Wu06}. This shape of element is a natural choice for a three-dimensional elastic body with a rectangular cross-section.

For the this three-dimensional model, piecewise Hermite tri-cubic basis functions are used. Although this is more complex than using piecwise tri-linear basis functions, the tri-cubic basis functions ensure faster convergence and also the derivatives of the solutions.

Now Problem 2DE and Problem 3DE can be solved using a computer program. Iterative methods, sparse matrices as well as parallel computing can be used to efficiently solve these problems.

\section{Validity of a model for a cantilever two-dimensional beam} \label{sec:validity-of-a-2d-beam}
As mentioned in the introduction of the chapter, a beam is a three-dimensional body and therefore a three-dimensional model is more realistic. However the authors of \cite{LVV09} mention that a direct comparison of the one and three-dimensional models will have complexities and they suggest using the two-dimensional model as an intermediate step.

@@

\subsection{The models}
Figure \ref{fig:2Dv3D} shows the two beams, Problem 2D and Problem 3D side-by-side.
